Analysis of Shootings and Firearm Discharges in Toronto

### 2024-12-02 paper/paper.qmd
reference update
paper update and parquet file added

The analysis employed a Gradient Boosting Machine (GBM) model[@gbm] using the `gbm` package in R, designed to predict the weighted scores of shootings and firearm discharges in Toronto based on multiple time-related and geographical predictors and using the `caret` package for advanced modeling techniques [@caret].

The weighted score is a computed metric that assigns double weight to deaths compared to injuries, highlighting the gravity of incidents resulting in fatalities.

@@

The response variable, weighted_score, combines the effects of deaths and injuries, assigning greater importance to the former. Predictors included:

@@

@@

• Number of Trees: 500, providing sufficient model complexity and accuracy.

• Interaction Depth: 4, allowing interactions among up to four predictors.

• Shrinkage: 0.01, controlling the learning rate to avoid overfitting.

• Cross-validation Folds: 5, used to validate the model internally and optimize parameter selection.

@@

@@

## Appendix: Advanced Methodologies for Sampling and Observational Data Analysis

## A1. Survey and Sampling Methodology

### A1.1 Overview

@@

### A1.2 Data Collection Method

@@

### A1.3 Sampling Method

@@

## A2. Addressing Biases and Limitations

### A2.1 Observational Bias

@@

### A2.2 Limitations of Observational Data

@@

## A3. Advanced Analytical Techniques

### A3.1 Model Specification

@@

### A3.2 Sensitivity Analysis

@@

### A3.3 Simulation Studies

@@

>>>>

The analysis employed a Gradient Boosting Machine (GBM) model(@gbm), utilizing the gbm package in R and caret package(@caret) to generate variable importance scores, which help in understanding which predictors are most influential in the model. This model type was selected for its robust performance with non-linear relationships and its ability to handle various types of data, which is crucial given the diverse nature of the dataset used. The GBM model excels in managing unbalanced data, such as the uneven distribution of shooting incidents across neighborhoods and times, and automatically handles interactions between variables, which is essential for our complex model.

@@

Response Variable:

The response variable, `weighted_score`, is a composite metric combining the effects of deaths and injuries, assigning greater importance to fatalities(`weighted_score` = `death` \* 2 + `injuries`). This decision was driven by the data characteristics where the distribution of shootings varied significantly by time and location, necessitating a response variable that captures the severity of incidents rather than mere counts.

Predictors included:

@@

Each predictor was chosen based on exploratory data analysis that indicated significant variations in shooting incidents associated with these variables. This choice ensures that the model can effectively learn and predict based on patterns specific to times, dates, and locations, which are critical in the dataset.

@@

**• Number of Trees**: 500, providing sufficient model complexity and accuracy.

**• Interaction Depth**: 4, allowing interactions among up to four predictors.

**• Shrinkage**: 0.01, controlling the learning rate to avoid overfitting.

**• Cross-validation Folds**: 5, used to validate the model internally and optimize parameter selection.

@@

### Linkage to Data Characteristics

The model’s configuration, especially the choice and treatment of predictor variables, directly corresponds to the observed characteristics of the data. For example, the use of `occ_time_range` and `neighbourhood_158` as factors allows the model to capture the inherent categorical nature of these variables, while numeric transformation of `occ_date` lets the model utilize temporal trends over the years. This direct linkage ensures that the model is finely tuned to the specifics of the dataset, enhancing its predictive accuracy and relevance.

@@

## Survey and Sampling Methodology

### Overview

@@

### Data Collection Method

@@

### Sampling Method

@@

## Addressing Biases and Limitations

### Observational Bias

@@

### Limitations of Observational Data

@@

## Advanced Analytical Techniques

### Model Specification

@@

### Sensitivity Analysis

@@

### Simulation Studies

@@

\newpage
